---
title: 'Chapter 1:  Visualization with ggplot'
author: "Augustina Ragwitz"
date: "June 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r imports}
library(tidyverse)
```

# First Steps

## Do cars with big engines use more fuel than cars with small engines?

### What is the relationship between engine-size and fuel efficiency?

This analysis uses the mpg dataset [^mpg].

```{r mpg}
summary(mpg)

paste("displ - the car's engine size, in liters:", paste(mpg$displ, collapse=","))

paste("hwy - the car's fuel efficiency on the highway, in mpg. Low fuel efficiency == more fuel consumed: ", 
      paste(mpg$hwy, collapse=","))
```


```{r fig-engine-size-vs-fuel-efficiency}

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))

```

The plot shows a negative relationship between engine size and fuel efficiency. A larger engine uses more fuel than a smaller engine.

## Exercises

### 1. Run `ggplot(data = mpg)`.

```{r exercise-1.1}
ggplot(data = mpg)
```

### 2. How many rows are in `mtcars`? How many columns?

```{r exercise-1.2}
paste("There are", nrow(mtcars), " rows in mtcars.")
paste("There are", ncol(mtcars), "columns in mtcars.")
```

### 3. What does the `drv` variable describe? Read `?mpg` to find out.

`drv` describes the "drive" or how many wheels on a car are engaged.

```{r exercise-1.3}
paste("drv: f = front-wheel drive, r = rear wheel drive, 4 = 4wd")
```


### 4. Make a scatterplot of `hwy` vs `cyl`.

```{r exercise-1.4}

ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))

```

The number of cylinders also appears to be related to the overall fuel-efficiency in some cases.

### 5. Make a scatterplot of `class` vs `drv`. Why isn't it useful?

There isn't anyway to see the frequency!

```{r exercise-1.5}

ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))

ggplot(data = mpg) +
  geom_bar(aes(x = class, fill = drv))

```

```{r exercise-1.5-better}

ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = class, y = drv, color = class, shape = drv))

```




---

[mpg]: "This dataset contains a subset of the fuel economy data that the EPA makes available on http://fueleconomy.gov. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car." From H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2009.